---
title: "Demographics" 
author: "Kendra Wyant"
date: "`r lubridate::today()`"
format: 
  html: 
    embed-resources: true
    toc: true 
    toc_depth: 4
editor_options: 
  chunk_output_type: console
---   

## Notes

This script calculates demographics for Burden2 study.

Outstanding issues:

- subid 1069 does not have intake survey.   
- Missing demographics for subid 1006  

- other things we might want to add to this tibble: dsm score and num lapses on study

## Set up
```{r}
#| message: false

options(conflicts.policy = "depends.ok")
library(tidyverse) 
library(janitor)

devtools::source_url("https://github.com/jjcurtin/lab_support/blob/main/format_path.R?raw=true")
devtools::source_url("https://github.com/jjcurtin/lab_support/blob/main/print_kbl.R?raw=true")

theme_set(theme_classic())

path_processed <- format_path("studydata/risk2/data_processed/shared")
path_burden <- format_path("studydata/risk2/data_processed/burden2")
```

## Read in data

Burden data
```{r}
burden <- read_csv(file.path(path_burden, "burden_quant.csv"),
                   show_col_types = FALSE)
```

Intake survey (with demographics)
```{r}
intake_survey <- read_csv(here::here(path_processed, "survey_intake.csv"),
                         show_col_types = FALSE) |> 
  filter(subid %in% burden$subid)
```



## Num intake surveys per subid
```{r}
length(unique(intake_survey$subid)) # number of unique subids (246)
length(unique(burden$subid)) # number of unique subids (247)
```

Subid 1069 does not have intake
```{r}
burden |> 
  filter(!subid %in% intake_survey$subid)
```


## Missing data
```{r}
intake_survey |> 
  filter(is.na(answer)) |> 
  janitor::tabyl(variable_name) |> 
  arrange(desc(n))
```

```{r}
intake_survey |> 
  filter(variable_name == "age" & is.na(answer))
```

Missing demographics for subid 1006
```{r}
intake_survey |> 
  filter(subid == 1006) |> 
  view()
```


## Create Demographics Tibble with variables of interest
```{r}
dem <- intake_survey |> 
  select(subid, variable_name, answer) |> 
  filter(variable_name %in% c("race", "gender", "orientation", "income",
                              "employment", "education", "age", "romantic_status",
                              "living_arrangement_1", "med", "whoassist_life_overdose")) 
```


Add missing demographics for subid 1069 (set to NA). 

```{r}
dem <- dem |> 
  bind_rows(tibble(subid = 1069,
                   variable_name = c("race", "gender", "orientation", "income",
                              "employment", "education", "age", "romantic_status",
                              "living_arrangement_1", "med", "whoassist_life_overdose"),
                   answer = NA_character_))
```

`r length(unique(dem$subid))` subids in final tibble.



## Summarize demographic categories  

### Race/ethnicity

```{r}
dem |> 
  filter(variable_name == "race") |> 
  tabyl(answer) |> 
  print_kbl()
```

Other 

```{r}
tabyl(subset(intake_survey, variable_name == "race_other")$answer)
```


Will create following race categories: White, Black, Hispanic, other

```{r}
dem |> 
  group_by()
```





### Gender

```{r}
dem |> 
  filter(variable_name == "gender") |> 
  tabyl(answer) |> 
  print_kbl()
```

### Orientation

```{r}
dem |> 
  filter(variable_name == "orientation") |> 
  tabyl(answer) |> 
  print_kbl()
```

### Income

```{r}
dem |> 
  filter(variable_name == "income") |> 
  tabyl(answer) |> 
  print_kbl()
```

### Employment

```{r}
dem |> 
  filter(variable_name == "employment") |> 
  tabyl(answer) |> 
  print_kbl()
```

### Education

```{r}
dem |> 
  filter(variable_name == "education") |> 
  tabyl(answer) |> 
  print_kbl()
```

### Age

```{r}
dem |> 
  filter(variable_name == "age") |> 
  tabyl(answer) |> 
  print_kbl()
```

### Committed Relationship

```{r}
dem |> 
  filter(variable_name == "romantic_status") |> 
  tabyl(answer) |> 
  print_kbl()
```

### Living Arrangement

```{r}
dem |> 
  filter(variable_name == "living_arrangement_1") |> 
  tabyl(answer) |> 
  print_kbl()
```

### Medication

```{r}
dem |> 
  filter(variable_name == "med") |> 
  tabyl(answer) |> 
  print_kbl()
```

### Lifetime History of Overdose

```{r}
dem |> 
  filter(variable_name == "whoassist_life_overdose") |> 
  tabyl(answer) |> 
  print_kbl()
```


Pivot wider and make label adjustments
```{r}
dem_final <- dem |> 
   pivot_wider(names_from = "variable_name", values_from = "answer") |> 
   mutate(race = if_else(race %in% c("Black/African American", 
                                    "Hispanic, Latino, or Spanish origin",
                                    "White/Caucasian", NA), race, "other"),
          gender = if_else(gender %in% c("Man", "Woman", NA), gender, "other"))
```


## Add location data


## Get total DSM score

```{r}
dsm <- intake_survey |> 
  filter(str_detect(variable_name, "dsm")) |> 
  select(subid, question_content, variable_name, answer) |> 
  mutate(answer = if_else(answer == "Yes", 1, 0))
```

```{r}
dsm <- dsm |> 
  group_by(subid) |> 
  mutate(dsm_criteria_1 = if_else((variable_name == "dsm_1" & answer == 1) | (variable_name == "dsm_2" & answer == 1) , 1, 0),
         dsm_criteria_2 = if_else(variable_name == "dsm_3" & answer == 1, 1, 0),
         dsm_criteria_3 = if_else((variable_name == "dsm_4" & answer == 1) | (variable_name == "dsm_5" & answer == 1) | (variable_name == "dsm_6" & answer == 1)  , 1, 0),
         dsm_criteria_4 = if_else((variable_name == "dsm_11" & answer == 1) , 1, 0),
         dsm_criteria_5 = if_else((variable_name == "dsm_7" & answer == 1) , 1, 0),
         dsm_criteria_6 = if_else((variable_name == "dsm_8" & answer == 1) , 1, 0),
         dsm_criteria_7 = if_else((variable_name == "dsm_9" & answer == 1) , 1, 0),
         dsm_criteria_8 = if_else((variable_name == "dsm_10" & answer == 1) , 1, 0),
         dsm_criteria_9 = if_else((variable_name == "dsm_12" & answer == 1) , 1, 0),
         dsm_criteria_10 = if_else((variable_name == "dsm_13" & answer == 1)  | (variable_name == "dsm_14" & answer == 1) , 1, 0),
         dsm_criteria_11 = if_else((variable_name == "dsm_15" & answer == 1)  | (variable_name == "dsm_16" & answer == 1) , 1, 0)) |> 
  summarise(dsm_criteria_1 = if_else(sum(dsm_criteria_1) >= 1, 1, 0),
            dsm_criteria_2 = if_else(sum(dsm_criteria_2) >= 1, 1, 0),
            dsm_criteria_3 = if_else(sum(dsm_criteria_3) >= 1, 1, 0),
            dsm_criteria_4 = if_else(sum(dsm_criteria_4) >= 1, 1, 0),
            dsm_criteria_5 = if_else(sum(dsm_criteria_5) >= 1, 1, 0),
            dsm_criteria_6 = if_else(sum(dsm_criteria_6) >= 1, 1, 0),
            dsm_criteria_7 = if_else(sum(dsm_criteria_7, na.rm = TRUE) >= 1, 1, 0),
            dsm_criteria_8 = if_else(sum(dsm_criteria_8, na.rm = TRUE) >= 1, 1, 0),
            dsm_criteria_9 = if_else(sum(dsm_criteria_9) >= 1, 1, 0),
            dsm_criteria_10 = if_else(sum(dsm_criteria_10) >= 1, 1, 0),
            dsm_criteria_11 = if_else(sum(dsm_criteria_10) >= 1, 1, 0)) 
```

```{r}
dsm <- dsm |> 
  rowwise() |> 
  mutate(dsm_total = sum(dsm_criteria_1, dsm_criteria_2, dsm_criteria_3,
                         dsm_criteria_4, dsm_criteria_5, dsm_criteria_6,
                         dsm_criteria_7, dsm_criteria_8, dsm_criteria_9,
                         dsm_criteria_10, dsm_criteria_11, na.rm = TRUE)) |> glimpse()
```

```{r}
dsm |> 
  ungroup() |> 
  summarize(min(dsm_total),
            max(dsm_total),
            mean(dsm_total),
            sd(dsm_total),
            median(dsm_total)) |> 
  glimpse()

tabyl(dsm$dsm_total)
```





## Save out tibble

```{r}
dem_final |> 
  write_csv(here::here(path_burden, "demographics.csv"))
```




